---
title: "4.3. PM and PE Models"
author: "Pablo Ruiz"
format: 
  html:
    self-contained: true
    toc: true
    toc-location: left
    toc-expand: 2
    theme: cosmo
editor: visual
---

# Load libraries and functions

```{r}
#| output: false

library(lubridate)
library(gam)
library(splines)
library(plyr)
library(flextable)
#source("0. Functions.R")

```

# 1. Load data

```{r}
load("Out/Bases2.RData")


```

# 2. Create time subsets

Usar este subset al final

```{r}

### CHECK DATES
summary(basesf_all_nse$fechaini)
summary(basesf_all_nse$fechafin)

### CREATE FINAL SUBSET (STRAND USING 2010-2016)
data<-basesf_all_nse[basesf_all_nse$fechafin>=as.Date("2010-01-01","%Y-%m-%d") &
                  basesf_all_nse$fechaini>=as.Date("2009-08-21","%Y-%m-%d") &
                  basesf_all_nse$fechaini<=as.Date("2016-03-05","%Y-%m-%d"), ]
nrow(data)
summary(data$fechaini)
summary(data$fechafin)

### CREATE DATA2 (STRIC STRAND)STRAND 2009-2016, POR QUÉ NO USAR 2009? DATA PM?)
data2<-basesf_all_nse[basesf_all_nse$fechafin>=as.Date("2009-01-01","%Y-%m-%d") &
                  basesf_all_nse$fechaini>=as.Date("2008-08-21","%Y-%m-%d") &
                  basesf_all_nse$fechaini<=as.Date("2016-03-05","%Y-%m-%d"), ]
nrow(data2)
summary(data2$fechaini)
summary(data2$fechafin)

```

# 3. Model preclampsi in base

## 3.1 Model sin cosin using bs

```{r}
#| warning: false

library(gam)
data<-as.data.frame(data)
data$time<-as.numeric(data$fechaini-as.Date("2009-01-01"))
model <- gam::gam(data$preclampsi ~ 
                s(data$time, df=3) +
                sin(2*pi/365*data$time) +
                cos(2*pi/365*data$time),
                family=binomial,
                na.action = na.exclude)

summary(model)

HacerLongSinMGCV<-function(data, var, time, df)
{
  # data
  data<-as.data.frame(data)
  data$var<-data[,var]
  data$time<-data[,time]
  data$time<-as.numeric(data$time-as.Date("2009-01-01"))
  
  # model
  model <- gam::gam(data$var ~ 
                  s(data$time, df=df) +
                  sin(2*pi/365*data$time) +
                  cos(2*pi/365*data$time),
                  family=binomial,
                  na.action = na.exclude)
  
  print(model)
  
  modelex<-gam.exact(model)
  print(modelex$coefficients)

  
  
  # plot model
  plot(fitted(model)~data$time,col=4,lty=2) 
 
  # plot Predict
  pr <- predict(model, type = "terms")
  print(summary(pr))
  print(paste("Suma:",sum(abs(pr[,2]+pr[,3]))))
  print(paste("Range:",max((pr[,2]+pr[,3])) - min((pr[,2]+pr[,3])) ))
  
    
  ### Plots
  par(mfrow = c(1, 1), mar = c(3, 4, 3, 2) )
  plot(pr[, 2]+pr[, 3] ~ data$time, ylab = "Trend", main = "long and sine cosine")
  points(pr[, 1] ~ data$time, ylab = "Trend", col="blue")
}


HacerLongSinMGCV(data, "preclampsi","fechaini",2)


```

# 4. Models Preeclampsia and PM

## 4.1 Flexible function

```{r}
#| warning: false

HacerModeloGAM<-function(data, 
                         outcome,
                         var, 
                         covs, 
                         covtime)
{
  ### DEAL DATA
  data<-as.data.frame(data)
  data$time<-as.numeric(data$fechaini-as.Date("2008-01-01"))
  
  
  ### FORMULA
  formula<-paste(outcome, " ~ ", var, " + ",covs," + ",covtime)
  print(formula)
  
  ### MODEL
  model <- gam::gam(as.formula(formula),
                  family=binomial,
                  na.action = na.exclude,
                  data=data)

  print(summary(model))
  
  ### PRINT COEFFICIENTS
  modelex<-gam.exact(model)
  print(round(modelex$coefficients,4))
}



HacerModeloGAMNoTime<-function(data, 
                         outcome,
                         var, 
                         covs)
{
  ### DEAL DATA
  data<-as.data.frame(data)
  data$time<-as.numeric(data$fechaini-as.Date("2008-01-01"))
  
  
  ### FORMULA
  formula<-paste(outcome, " ~ ", var, " + ",covs)
  print(formula)
  
  ### MODEL
  model <- gam::gam(as.formula(formula),
                  family=binomial,
                  na.action = na.exclude,
                  data=data)

  print(summary(model))
  
  ### PRINT COEFFICIENTS
  modelex<-gam.exact(model)
  print(round(modelex$coefficients,4))
}

  
```

## 4.2 Modelos nulo, age (s=1) + time (s=3) sin cosine

```{r}
#| warning: false

library(gam)
data<-as.data.frame(data)
data$time<-as.numeric(data$fechaini-as.Date("2008-01-01"))


### SIN PM
model <- gam::gam(data$preclampsi ~ s(data$age2, df = 1) + 
                sin(2*pi/365*data$time) +
                cos(2*pi/365*data$time),
                family=binomial,
                na.action = na.exclude)

summary(model)

modelex<-gam.exact(model)
print(modelex$coefficients)

### REDO WITH FUNCTION
HacerModeloGAM(data=data, 
                outcome="preclampsi",
                var="s(age2, df = 1)", 
                covs="", 
                covtime="sin(2*pi/365*time) + cos(2*pi/365*time)")


```

## 4.3 Modelos con PM + age(s=1) solamente

```{r}

#### Central Site CS

HacerModeloGAMNoTime(data=data, 
                outcome="preclampsi",
                var="pc24w_PM25_cs", 
                covs="s(age2, df = 1)")

HacerModeloGAMNoTime(data=data, 
                outcome="preclampsi",
                var="pc12w_PM25_cs", 
                covs="s(age2, df = 1)")

HacerModeloGAMNoTime(data=data, 
                outcome="preclampsi",
                var="t1_PM25_cs", 
                covs="s(age2, df = 1)")

HacerModeloGAMNoTime(data=data, 
                outcome="preclampsi",
                var="t2_PM25_cs", 
                covs="s(age2, df = 1)")

HacerModeloGAMNoTime(data=data, 
                outcome="preclampsi",
                var="t3_PM25_cs", 
                covs="s(age2, df = 1)")

HacerModeloGAMNoTime(data=data, 
                outcome="preclampsi",
                var="w20_PM25_cs", 
                covs="s(age2, df = 1)")

HacerModeloGAMNoTime(data=data, 
                outcome="preclampsi",
                var="tot_PM25_cs", 
                covs="s(age2, df = 1)")

HacerModeloGAMNoTime(data=data, 
                outcome="preclampsi",
                var="t1_PM25_cs + t2_PM25_cs + t3_PM25_cs", 
                covs="s(age2, df = 1)")

HacerModeloGAMNoTime(data=data, 
                outcome="preclampsi",
                var="pc24w_PM25_cs + t1_PM25_cs + t2_PM25_cs + t3_PM25_cs", 
                covs="s(age2, df = 1)")


#### SPATIAL LUR SP

HacerModeloGAMNoTime(data=data, 
                outcome="preclampsi",
                var="pc24w_PM25_sp", 
                covs="s(age2, df = 1)")

HacerModeloGAMNoTime(data=data, 
                outcome="preclampsi",
                var="pc12w_PM25_sp", 
                covs="s(age2, df = 1)")

HacerModeloGAMNoTime(data=data, 
                outcome="preclampsi",
                var="t1_PM25_sp", 
                covs="s(age2, df = 1)")

HacerModeloGAMNoTime(data=data, 
                outcome="preclampsi",
                var="t2_PM25_sp", 
                covs="s(age2, df = 1)")

HacerModeloGAMNoTime(data=data, 
                outcome="preclampsi",
                var="t3_PM25_sp", 
                covs="s(age2, df = 1)")

HacerModeloGAMNoTime(data=data, 
                outcome="preclampsi",
                var="w20_PM25_sp", 
                covs="s(age2, df = 1)")

HacerModeloGAMNoTime(data=data, 
                outcome="preclampsi",
                var="tot_PM25_sp", 
                covs="s(age2, df = 1)")

HacerModeloGAMNoTime(data=data, 
                outcome="preclampsi",
                var="t1_PM25_sp + t2_PM25_sp + t3_PM25_sp", 
                covs="s(age2, df = 1)")

HacerModeloGAMNoTime(data=data, 
                outcome="preclampsi",
                var="pc24w_PM25_sp + t1_PM25_sp + t2_PM25_sp + t3_PM25_sp", 
                covs="s(age2, df = 1)")

```

## 4.4 Modelos PM y age (s=1) + time (var df) + sin cosine

### 4.4.1. t1

```{r}


HacerModeloGAM(data=data, 
                outcome="preclampsi",
                var="t1_PM25_cs", 
                covs="s(age2, df = 1)",
                covtime="s(data$time, df=1) +
                sin(2*pi/365*data$time) +
                cos(2*pi/365*data$time)")

HacerModeloGAM(data=data, 
                outcome="preclampsi",
                var="t1_PM25_cs", 
                covs="s(age2, df = 1)",
                covtime="s(data$time, df=2) +
                sin(2*pi/365*data$time) +
                cos(2*pi/365*data$time)")

HacerModeloGAM(data=data, 
                outcome="preclampsi",
                var="t1_PM25_cs", 
                covs="s(age2, df = 1)",
                covtime="s(data$time, df=3) +
                sin(2*pi/365*data$time) +
                cos(2*pi/365*data$time)")

HacerModeloGAM(data=data, 
                outcome="preclampsi",
                var="t1_PM25_cs", 
                covs="s(age2, df = 1)",
                covtime="s(data$time, df=4) +
                sin(2*pi/365*data$time) +
                cos(2*pi/365*data$time)")
```

### 4.4.2. t2

```{r}

HacerModeloGAM(data=data, 
                outcome="preclampsi",
                var="t2_PM25_cs", 
                covs="s(age2, df = 1)",
                covtime="s(data$time, df=1) +
                sin(2*pi/365*data$time) +
                cos(2*pi/365*data$time)")

HacerModeloGAM(data=data, 
                outcome="preclampsi",
                var="t2_PM25_cs", 
                covs="s(age2, df = 1)",
                covtime="s(data$time, df=2) +
                sin(2*pi/365*data$time) +
                cos(2*pi/365*data$time)")

HacerModeloGAM(data=data, 
                outcome="preclampsi",
                var="t2_PM25_cs", 
                covs="s(age2, df = 1)",
                covtime="s(data$time, df=3) +
                sin(2*pi/365*data$time) +
                cos(2*pi/365*data$time)")

HacerModeloGAM(data=data, 
                outcome="preclampsi",
                var="t2_PM25_cs", 
                covs="s(age2, df = 1)",
                covtime="s(data$time, df=4) +
                sin(2*pi/365*data$time) +
                cos(2*pi/365*data$time)")

HacerModeloGAM(data=data, 
                outcome="preclampsi",
                var="t2_PM25_cs", 
                covs="s(age2, df = 1)",
                covtime="s(data$time, df=5) +
                sin(2*pi/365*data$time) +
                cos(2*pi/365*data$time)")


```

## 4.5. PM (vary) + age(s=1) + time(s=6df)

```{r}

HacerModeloGAM(data=data, 
                outcome="preclampsi",
                var="pc24w_PM25_cs", 
                covs="s(age2, df = 1)",
                covtime="s(data$time, df=6)")

HacerModeloGAM(data=data, 
                outcome="preclampsi",
                var="pc12w_PM25_cs", 
                covs="s(age2, df = 1)",
                covtime="s(data$time, df=6)")

HacerModeloGAM(data=data, 
                outcome="preclampsi",
                var="t1_PM25_cs", 
                covs="s(age2, df = 1)",
                covtime="s(data$time, df=6)")

HacerModeloGAM(data=data, 
                outcome="preclampsi",
                var="t2_PM25_cs", 
                covs="s(age2, df = 1)",
                covtime="s(data$time, df=6)")

HacerModeloGAM(data=data, 
                outcome="preclampsi",
                var="t3_PM25_cs", 
                covs="s(age2, df = 1)",
                covtime="s(data$time, df=6)")

HacerModeloGAM(data=data, 
                outcome="preclampsi",
                var="w20_PM25_cs", 
                covs="s(age2, df = 1)",
                covtime="s(data$time, df=6)")


HacerModeloGAM(data=data, 
                outcome="preclampsi",
                var="tot_PM25_cs", 
                covs="s(age2, df = 1)",
                covtime="s(data$time, df=6)")

HacerModeloGAM(data=data, 
                outcome="preclampsi",
                var="t1_PM25_cs + t2_PM25_cs + t3_PM25_cs", 
                covs="s(age2, df = 1)",
                covtime="s(data$time, df=6)")

HacerModeloGAM(data=data, 
                outcome="preclampsi",
                var="pc24w_PM25_cs + t1_PM25_cs + t2_PM25_cs + t3_PM25_cs", 
                covs="s(age2, df = 1)",
                covtime="s(data$time, df=6)")

HacerModeloGAM(data=data, 
                outcome="preclampsi",
                var="pc24w_PM25_cs + t2_PM25_cs", 
                covs="s(age2, df = 1)",
                covtime="s(data$time, df=6)")

HacerModeloGAM(data=data, 
                outcome="preclampsi",
                var="pc12w_PM25_cs + t2_PM25_cs", 
                covs="s(age2, df = 1)",
                covtime="s(data$time, df=6)")


```

## 4.6. PM (vary) + age(s=1) + time(s=var) + sincos

### 4.6.1. time s=4df

```{r}

HacerModeloGAM(data=data, 
                outcome="preclampsi",
                var="pc24w_PM25_cs", 
                covs="s(age2, df = 1)",
                covtime="s(data$time, df=4) +
                sin(2*pi/365*data$time) +
                cos(2*pi/365*data$time)")

HacerModeloGAM(data=data, 
                outcome="preclampsi",
                var="pc12w_PM25_cs", 
                covs="s(age2, df = 1)",
                covtime="s(data$time, df=4) +
                sin(2*pi/365*data$time) +
                cos(2*pi/365*data$time)")

HacerModeloGAM(data=data, 
                outcome="preclampsi",
                var="t1_PM25_cs", 
                covs="s(age2, df = 1)",
                covtime="s(data$time, df=4) +
                sin(2*pi/365*data$time) +
                cos(2*pi/365*data$time)")

HacerModeloGAM(data=data, 
                outcome="preclampsi",
                var="t2_PM25_cs", 
                covs="s(age2, df = 1)",
                covtime="s(data$time, df=4) +
                sin(2*pi/365*data$time) +
                cos(2*pi/365*data$time)")

HacerModeloGAM(data=data, 
                outcome="preclampsi",
                var="t3_PM25_cs", 
                covs="s(age2, df = 1)",
                covtime="s(data$time, df=4) +
                sin(2*pi/365*data$time) +
                cos(2*pi/365*data$time)")

HacerModeloGAM(data=data, 
                outcome="preclampsi",
                var="w20_PM25_cs", 
                covs="s(age2, df = 1)",
                covtime="s(data$time, df=4) +
                sin(2*pi/365*data$time) +
                cos(2*pi/365*data$time)")


HacerModeloGAM(data=data, 
                outcome="preclampsi",
                var="tot_PM25_cs", 
                covs="s(age2, df = 1)",
                covtime="s(data$time, df=4) +
                sin(2*pi/365*data$time) +
                cos(2*pi/365*data$time)")

HacerModeloGAM(data=data, 
                outcome="preclampsi",
                var="t1_PM25_cs + t2_PM25_cs + t3_PM25_cs", 
                covs="s(age2, df = 1)",
                covtime="s(data$time, df=4) +
                sin(2*pi/365*data$time) +
                cos(2*pi/365*data$time)")

HacerModeloGAM(data=data, 
                outcome="preclampsi",
                var="pc12w_PM25_cs + t1_PM25_cs + t2_PM25_cs + t3_PM25_cs", 
                covs="s(age2, df = 1)",
                covtime="s(data$time, df=4) +
                sin(2*pi/365*data$time) +
                cos(2*pi/365*data$time)")

HacerModeloGAM(data=data, 
                outcome="preclampsi",
                var="pc24w_PM25_cs + t1_PM25_cs + t2_PM25_cs + t3_PM25_cs", 
                covs="s(age2, df = 1)",
                covtime="s(data$time, df=4) +
                sin(2*pi/365*data$time) +
                cos(2*pi/365*data$time)")

```

### 4.6.2. cs time s=6df

 [ **Best models**]{style="color:red; font-size: 200%;"}
```{r}

HacerModeloGAM(data=data, 
                outcome="preclampsi",
                var="pc24w_PM25_cs", 
                covs="s(age2, df = 1)",
                covtime="s(data$time, df=6) +
                sin(2*pi/365*data$time) +
                cos(2*pi/365*data$time)")

HacerModeloGAM(data=data, 
                outcome="preclampsi",
                var="pc12w_PM25_cs", 
                covs="s(age2, df = 1)",
                covtime="s(data$time, df=6) +
                sin(2*pi/365*data$time) +
                cos(2*pi/365*data$time)")

HacerModeloGAM(data=data, 
                outcome="preclampsi",
                var="t1_PM25_cs", 
                covs="s(age2, df = 1)",
                covtime="s(data$time, df=6) +
                sin(2*pi/365*data$time) +
                cos(2*pi/365*data$time)")

HacerModeloGAM(data=data, 
                outcome="preclampsi",
                var="t2_PM25_cs", 
                covs="s(age2, df = 1)",
                covtime="s(data$time, df=6) +
                sin(2*pi/365*data$time) +
                cos(2*pi/365*data$time)")

HacerModeloGAM(data=data, 
                outcome="preclampsi",
                var="t3_PM25_cs", 
                covs="s(age2, df = 1)",
                covtime="s(data$time, df=6) +
                sin(2*pi/365*data$time) +
                cos(2*pi/365*data$time)")

HacerModeloGAM(data=data, 
                outcome="preclampsi",
                var="w20_PM25_cs", 
                covs="s(age2, df = 1)",
                covtime="s(data$time, df=6) +
                sin(2*pi/365*data$time) +
                cos(2*pi/365*data$time)")


HacerModeloGAM(data=data, 
                outcome="preclampsi",
                var="tot_PM25_cs", 
                covs="s(age2, df = 1)",
                covtime="s(data$time, df=6) +
                sin(2*pi/365*data$time) +
                cos(2*pi/365*data$time)")

HacerModeloGAM(data=data, 
                outcome="preclampsi",
                var="t1_PM25_cs + t2_PM25_cs + t3_PM25_cs", 
                covs="s(age2, df = 1)",
                covtime="s(data$time, df=6) +
                sin(2*pi/365*data$time) +
                cos(2*pi/365*data$time)")

HacerModeloGAM(data=data, 
                outcome="preclampsi",
                var="pc24w_PM25_cs + t1_PM25_cs + t2_PM25_cs + t3_PM25_cs", 
                covs="s(age2, df = 1)",
                covtime="s(data$time, df=6) +
                sin(2*pi/365*data$time) +
                cos(2*pi/365*data$time)")

HacerModeloGAM(data=data, 
                outcome="preclampsi",
                var="pc12w_PM25_cs + t1_PM25_cs + t2_PM25_cs + t3_PM25_cs", 
                covs="s(age2, df = 1)",
                covtime="s(data$time, df=6) +
                sin(2*pi/365*data$time) +
                cos(2*pi/365*data$time)")


### GOOD
HacerModeloGAM(data=data, 
                outcome="preclampsi",
                var="pc24w_PM25_cs + t2_PM25_cs", 
                covs="s(age2, df = 1)",
                covtime="s(data$time, df=6) +
                sin(2*pi/365*data$time) +
                cos(2*pi/365*data$time)")


HacerModeloGAM(data=data, 
                outcome="preclampsi",
                var="pc12w_PM25_cs + t2_PM25_cs", 
                covs="s(age2, df = 1)",
                covtime="s(data$time, df=6) +
                sin(2*pi/365*data$time) +
                cos(2*pi/365*data$time)")

```

### 4.6.3. sp time s=6df

```{r}

HacerModeloGAM(data=data, 
                outcome="preclampsi",
                var="pc24w_PM25_sp", 
                covs="s(age2, df = 1)",
                covtime="s(data$time, df=6) +
                sin(2*pi/365*data$time) +
                cos(2*pi/365*data$time)")

HacerModeloGAM(data=data, 
                outcome="preclampsi",
                var="pc12w_PM25_sp", 
                covs="s(age2, df = 1)",
                covtime="s(data$time, df=6) +
                sin(2*pi/365*data$time) +
                cos(2*pi/365*data$time)")

HacerModeloGAM(data=data, 
                outcome="preclampsi",
                var="t1_PM25_sp", 
                covs="s(age2, df = 1)",
                covtime="s(data$time, df=6) +
                sin(2*pi/365*data$time) +
                cos(2*pi/365*data$time)")

HacerModeloGAM(data=data, 
                outcome="preclampsi",
                var="t2_PM25_sp", 
                covs="s(age2, df = 1)",
                covtime="s(data$time, df=6) +
                sin(2*pi/365*data$time) +
                cos(2*pi/365*data$time)")

HacerModeloGAM(data=data, 
                outcome="preclampsi",
                var="t3_PM25_sp", 
                covs="s(age2, df = 1)",
                covtime="s(data$time, df=6) +
                sin(2*pi/365*data$time) +
                cos(2*pi/365*data$time)")

HacerModeloGAM(data=data, 
                outcome="preclampsi",
                var="w20_PM25_sp", 
                covs="s(age2, df = 1)",
                covtime="s(data$time, df=6) +
                sin(2*pi/365*data$time) +
                cos(2*pi/365*data$time)")


HacerModeloGAM(data=data, 
                outcome="preclampsi",
                var="tot_PM25_sp", 
                covs="s(age2, df = 1)",
                covtime="s(data$time, df=6) +
                sin(2*pi/365*data$time) +
                cos(2*pi/365*data$time)")

HacerModeloGAM(data=data, 
                outcome="preclampsi",
                var="t1_PM25_sp + t2_PM25_sp + t3_PM25_sp", 
                covs="s(age2, df = 1)",
                covtime="s(data$time, df=6) +
                sin(2*pi/365*data$time) +
                cos(2*pi/365*data$time)")

HacerModeloGAM(data=data, 
                outcome="preclampsi",
                var="pc24w_PM25_sp + t1_PM25_sp + t2_PM25_sp + t3_PM25_sp", 
                covs="s(age2, df = 1)",
                covtime="s(data$time, df=6) +
                sin(2*pi/365*data$time) +
                cos(2*pi/365*data$time)")

HacerModeloGAM(data=data, 
                outcome="preclampsi",
                var="pc12w_PM25_sp + t1_PM25_sp + t2_PM25_sp + t3_PM25_sp", 
                covs="s(age2, df = 1)",
                covtime="s(data$time, df=6) +
                sin(2*pi/365*data$time) +
                cos(2*pi/365*data$time)")


### GOOD
HacerModeloGAM(data=data, 
                outcome="preclampsi",
                var="pc24w_PM25_sp + t2_PM25_sp", 
                covs="s(age2, df = 1)",
                covtime="s(data$time, df=6) +
                sin(2*pi/365*data$time) +
                cos(2*pi/365*data$time)")


HacerModeloGAM(data=data, 
                outcome="preclampsi",
                var="pc12w_PM25_sp + t2_PM25_sp", 
                covs="s(age2, df = 1)",
                covtime="s(data$time, df=6) +
                sin(2*pi/365*data$time) +
                cos(2*pi/365*data$time)")

```

### 4.6.4. cs time s=6df sincos 4x

```{r}

HacerModeloGAM(data=data, 
                outcome="preclampsi",
                var="pc24w_PM25_cs", 
                covs="s(age2, df = 1)",
                covtime="s(data$time, df=6) +
                sin(2*pi/365*data$time) +
                cos(2*pi/365*data$time) +
                sin(4*pi/365*data$time) +
                cos(4*pi/365*data$time) ")

HacerModeloGAM(data=data, 
                outcome="preclampsi",
                var="pc12w_PM25_cs", 
                covs="s(age2, df = 1)",
                covtime="s(data$time, df=6) +
                sin(2*pi/365*data$time) +
                cos(2*pi/365*data$time) +
                sin(4*pi/365*data$time) +
                cos(4*pi/365*data$time)")

HacerModeloGAM(data=data, 
                outcome="preclampsi",
                var="t1_PM25_cs", 
                covs="s(age2, df = 1)",
                covtime="s(data$time, df=6) +
                sin(2*pi/365*data$time) +
                cos(2*pi/365*data$time) +
                sin(4*pi/365*data$time) +
                cos(4*pi/365*data$time)")

HacerModeloGAM(data=data, 
                outcome="preclampsi",
                var="t2_PM25_cs", 
                covs="s(age2, df = 1)",
                covtime="s(data$time, df=6) +
                sin(2*pi/365*data$time) +
                cos(2*pi/365*data$time) +
                sin(4*pi/365*data$time) +
                cos(4*pi/365*data$time)")

HacerModeloGAM(data=data, 
                outcome="preclampsi",
                var="t3_PM25_cs", 
                covs="s(age2, df = 1)",
                covtime="s(data$time, df=6) +
                sin(2*pi/365*data$time) +
                cos(2*pi/365*data$time) +
                sin(4*pi/365*data$time) +
                cos(4*pi/365*data$time)")

HacerModeloGAM(data=data, 
                outcome="preclampsi",
                var="w20_PM25_cs", 
                covs="s(age2, df = 1)",
                covtime="s(data$time, df=6) +
                sin(2*pi/365*data$time) +
                cos(2*pi/365*data$time) +
                sin(4*pi/365*data$time) +
                cos(4*pi/365*data$time)")


HacerModeloGAM(data=data, 
                outcome="preclampsi",
                var="tot_PM25_cs", 
                covs="s(age2, df = 1)",
                covtime="s(data$time, df=6) +
                sin(2*pi/365*data$time) +
                cos(2*pi/365*data$time) +
                sin(4*pi/365*data$time) +
                cos(4*pi/365*data$time)")

HacerModeloGAM(data=data, 
                outcome="preclampsi",
                var="t1_PM25_cs + t2_PM25_cs + t3_PM25_cs", 
                covs="s(age2, df = 1)",
                covtime="s(data$time, df=6) +
                sin(2*pi/365*data$time) +
                cos(2*pi/365*data$time) +
                sin(4*pi/365*data$time) +
                cos(4*pi/365*data$time)")

HacerModeloGAM(data=data, 
                outcome="preclampsi",
                var="pc24w_PM25_cs + t1_PM25_cs + t2_PM25_cs + t3_PM25_cs", 
                covs="s(age2, df = 1)",
                covtime="s(data$time, df=6) +
                sin(2*pi/365*data$time) +
                cos(2*pi/365*data$time) +
                sin(4*pi/365*data$time) +
                cos(4*pi/365*data$time)")

HacerModeloGAM(data=data, 
                outcome="preclampsi",
                var="pc12w_PM25_cs + t1_PM25_cs + t2_PM25_cs + t3_PM25_cs", 
                covs="s(age2, df = 1)",
                covtime="s(data$time, df=6) +
                sin(2*pi/365*data$time) +
                cos(2*pi/365*data$time) +
                sin(4*pi/365*data$time) +
                cos(4*pi/365*data$time)")


### GOOD
HacerModeloGAM(data=data, 
                outcome="preclampsi",
                var="pc24w_PM25_cs + t2_PM25_cs", 
                covs="s(age2, df = 1)",
                covtime="s(data$time, df=6) +
                sin(2*pi/365*data$time) +
                cos(2*pi/365*data$time) +
                sin(4*pi/365*data$time) +
                cos(4*pi/365*data$time)")


HacerModeloGAM(data=data, 
                outcome="preclampsi",
                var="pc12w_PM25_cs + t2_PM25_cs", 
                covs="s(age2, df = 1)",
                covtime="s(data$time, df=6) +
                sin(2*pi/365*data$time) +
                cos(2*pi/365*data$time) +
                sin(4*pi/365*data$time) +
                cos(4*pi/365*data$time)")

```
